---
title: "The 'Mapping Birdsong' Project"
author: "Mike Allen"
date: "1/31/2022"
output: html_document
---
# Load libraries and read in your data file
This code file is organized into discrete "chunks" each of which can be run by clicking its green arrow. Make sure there is only one file (the csv you downloaded from Macaulay Library) in the "my_raw_data" folder. You'll need to delete the Prairie Warbler example file that is currently there. This code reads in the data, keeping only songs with a known location (latitude/longitude) and with a quality score of 4 or greater.
```{r}
# load the function libraries called "dplyr" and "ggplot2"
# note: this also installs them if they are not yet installed
if (require("dplyr") == F) {
  install.packages("dplyr")
}
if (require("ggplot2") == F) {
  install.packages("ggplot2")
}

library(dplyr)
library(ggplot2)

# read in the data from your "my_raw_data" folder
# note: it also prints a warning if you have >1 file in your folder
if (length(list.files("my_raw_data")) != 1) {
  warning(
    "Only 1 file can be in the my_raw_data folder. Please delete any except the one you want to analyze."
  )
} else{
  my_data <- read.csv(paste0("my_raw_data/",
                             list.files("my_raw_data"))) %>%
    filter(Average.Community.Rating >= 4,
           is.na(Latitude) != TRUE) %>%
    rename(cat_num = 1)
}
```
# Calculate how many samples to take from each latitude band
Hit the green arrow to run this code that determines how many songs to randomly select from each latitude band so that we end up with about 25-50 songs spread out across the range of the species.
```{r}
# calculate number of latitude bands your data spans
  num_lat_bands <-
    length(unique(floor(my_data$Latitude)))

# calculate number of latitude bands your data spans
  num_lon_bands <-
    length(unique(floor(my_data$Longitude)))

# calculate the number of samples to take from each band
if(num_lat_bands >= 25){num_lat_samps <- 1}
if(num_lat_bands < 25){num_lat_samps <- 2}
if(num_lat_bands < 12){num_lat_samps <- 3}
if(num_lon_bands >= 25){num_lon_samps <- 1}
if(num_lon_bands < 25){num_lon_samps <- 2}
if(num_lon_bands < 12){num_lon_samps <- 3}
```
# Create final the data file
Hit the green arrow to run this code that creates two final data files, each with about 25-50 songs spread out across the species range (by latitude or longitude, respectively) that you will use for taking measurements.
```{r}
# create a data set stratified by latitude: my_data_final_lat
my_data_final_lat <- my_data %>%
  arrange(runif(1:nrow(my_data))) %>%
  group_by(floor(Latitude)) %>%
  mutate(nbin = length(floor(Latitude))) %>%
  filter(nbin >= num_lat_samps) %>%
  slice_sample(n = num_lat_samps) %>%
  ungroup() %>%
  select(cat_num, Latitude, Longitude) %>%
  mutate(link = paste0("https://macaulaylibrary.org/asset/", 
                       cat_num),
         measurement = "", measurement_units = "", 
         skip_notes = ""
)

# create a data set stratified by longitude: my_data_final_lon
my_data_final_lon <- my_data %>%
  arrange(runif(1:nrow(my_data))) %>%
  group_by(floor(Longitude)) %>%
  mutate(nbin = length(floor(Longitude))) %>%
  filter(nbin >= num_lon_samps) %>%
  slice_sample(n = num_lon_samps) %>%
  ungroup() %>%
  select(cat_num, Latitude, Longitude) %>%
  mutate(link = paste0("https://macaulaylibrary.org/asset/", 
                       cat_num),
         measurement = "", measurement_units = "", 
         skip_notes = ""
)

# write the final "blank" data files to csv files

write.csv(my_data_final_lat, 
          "my_output/my_blank_data_file_latitude.csv",
          row.names = F)

write.csv(my_data_final_lon, 
          "my_output/my_blank_data_file_longitude.csv",
          row.names = F)

```
# Making your measurements
At this point, minimize RStudio, go back to the README file, and complete steps 9-14. In those steps, you'll use spreadsheet software to open the "my_data_file.csv" file you just created. Then, you'll go to each link, view the sonogram, and make the measurements you came up with using a ruler or visual observation. (See README file for more info.) You'll add those measurements to your data file and then come back here to graph the results.

# Graph the results
Use this code to plot your results. It will save a png file to the "my_output" folder. Before you run it, change the 2nd and 3rd lines of the code to reflect your correct units and species. After you run this chunk of code, go ahead and complete the remaining sections of code in this file (described in README steps 15-17).
# First input the correct measurement units and species
Change the quoted text to reflect your own study.
```{r}
your_units <- "Number of syllables in trill"
your_species <- "Prairie Warbler"
```
# Make the plots
Run this chunk (click the green arrow) to make and save the plots. They should appear in your "my_output".
```{r}
# check if output files exist
lat_exists <- 
  file.exists("my_output/my_complete_data_file_latitude.csv")
lon_exists <- 
  file.exists("my_output/my_complete_data_file_longitude.csv")

# read in the completed latitude data file
if(lat_exists){
data_complete_lat <- read.csv(
  "my_output/my_complete_data_file_latitude.csv"
  )
}

# read in the completed longitude data file
if(lon_exists){
data_complete_lon <- read.csv(
  "my_output/my_complete_data_file_longitude.csv"
  )
}

# combine the two data files (if they exist)
if(lat_exists & lon_exists){
  data_complete <- 
    data_complete_lat %>%
    bind_rows(data_complete_lon)
}

if(lat_exists==T & lon_exists==F){
  data_complete <- data_complete_lat}

if(lat_exists==F & lon_exists==T){
  data_complete <- data_complete_lon}

# make sure you have the required packages just in case
if(require(ggplot2)==F){install.packages("ggplot2")
  library(ggplot2)}
if(require(dplyr)==F){install.packages("dplyr")
  library(dplyr)}

# plot the measurement by latitude
data_complete %>%
  ggplot(aes(x = Latitude, y = measurement)) +
  geom_point(size = 4) +
  geom_smooth(method = "loess", se = F, color = "firebrick") +
  geom_smooth(method = "lm", se = T, color = "darkgray", lty = 2) +
  theme_bw() + 
  theme(text = element_text(size = 14)) +
  labs(y = your_units, title = your_species)

# save the latitude plot
ggsave("my_output/my_plot_latitude.png", 
       height = 4, width = 6, dpi = 100)

# plot the measurement by longitude
data_complete %>%
  ggplot(aes(x = Longitude, y = measurement)) +
  geom_point(size = 4) +
  geom_smooth(method = "loess", se = F, color = "firebrick") +
  geom_smooth(method = "lm", se = T, color = "darkgray", lty = 2) +
  theme_bw() + 
  theme(text = element_text(size = 14)) +
  labs(y = your_units, title = your_species)

# save the longitude plot
ggsave("my_output/my_plot_longitude.png", 
       height = 4, width = 6, dpi = 100)

```
# Map the results
```{r}
# make sure you have the required packages installed
if(require(rnaturalearth)==F){install.packages("rnaturalearth")}
if(require(sf)==F){install.packages("sf")}
if(require(viridis)==F){install.packages("viridis")}
if(require(ggplot2)==F){install.packages("ggplot2")}
if(require(dplyr)==F){install.packages("dplyr")}
library(rnaturalearth)
library(sf)
library(viridis)
library(ggplot2)
library(dplyr)

# make the data into a "spatial" format
data_spatial <- data_complete %>%
  mutate(lat = Latitude, lon = Longitude) %>%
  sf::st_as_sf(coords = c("Longitude", "Latitude"), 
               crs = 4326) 

# load the continental boundaries
state_bounds <- ne_coastline(scale = 10, returnclass = "sf") %>%
  st_crop(st_buffer(data_spatial, dist = 2000000))

# make the map
data_spatial %>%
  ggplot() +
  geom_point(aes(x = lon, y = lat, 
                 size = measurement,
                 color = measurement), 
            alpha = .75) +
  viridis::scale_color_viridis(option = "magma") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) + 
  theme_bw() + 
  theme(text = element_text(size = 14)) +
  labs(size = your_units, title = your_species,
       x = "", y = "", color = your_units)

# save the map
ggsave("my_output/my_map.png", 
       height = 6, width = 7, dpi = 100)

```
# Describe your methods
Replace the DESCRIBE YOUR METHODS HERE text with your own. Don't delete any of the quotation marks or other text. Be sure to clearly describe what you measured and how. Write as much as you need to, but only replace the text within the asterisks(*). After you are done writing, run the code to write a text file using the green arrow.
```{r}
cat(file = "my_output/my_methods.txt", 
    "
METHODS

*DESCRIBE YOUR METHODS HERE*


    "
)

```
# Describe your results
Replace the DESCRIBE YOUR RESULTS HERE text with your own. Don't delete any of the quotation marks or other text. Write as much as you feel you need to, but only replace the text within the asterisks(*). After you are done writing, run the code to write a text file using the green arrow.
```{r}
cat(file = "my_output/my_results.txt", 
    "
RESULTS

*DESCRIBE YOUR RESULTS HERE*


    "
)
```
