---
title: "Mapping Birdsong - Latitude"
author: "Mike Allen"
date: "1/31/2022"
output: html_document
---
# Load libraries and read in your data file
Click the green arrow to run this code. Make sure there is only one file (the csv you downloaded from Macaulay Library) in the "my_raw_data" folder. Note that it only keeps songs with a known location (lattitude) and with a quality score of 4 or greater.
```{r}
# load the function libraries called "dplyr" and "ggplot2"
# note: this also installs them if they are not yet installed
if (require("dplyr") == F) {
  install.packages("dplyr")
}
if (require("ggplot2") == F) {
  install.packages("ggplot2")
}

# read in the data from your "my_raw_data" folder
# note: it also prints a warning if you have >1 file in your folder
if (length(list.files("my_raw_data")) != 1) {
  warning(
    "Only 1 file can be in the my_raw_data folder. Please delete any except the one you want to analyze."
  )
} else{
  my_data <- read.csv(paste0("my_raw_data/",
                             list.files("my_raw_data"))) %>%
    filter(Average.Community.Rating >= 4,
           is.na(Latitude) != TRUE) %>%
    rename(cat_num = 1)
}
```
# Calculate how many samples to take from each latitude band
Hit the green arrow to run this code that determines how many songs to randomly select from each latitude band so that we end up with about 25-50 songs spread out across the range of the species.
```{r}
# calculate number of latitude bands your data spans
  num_bands <-
    length(unique(floor(my_data$Latitude)))

# calculate the number of samples to take from each band
if(num_bands >= 25){num_samps <- 1}
if(num_bands < 25){num_samps <- 2}
if(num_bands < 12){num_samps <- 3}
```
# Create final the data file
Hit the green arrow to run this code that creates a final data file with about 25-50 songs spread out across the species range that you will use for taking measurements.
```{r}
# create a data set called my_data_final
my_data_final <- my_data %>%
  arrange(runif(1:nrow(my_data))) %>%
  group_by(floor(Latitude)) %>%
  slice_sample(n = num_samps) %>%
  select(cat_num, Latitude, Longitude) %>%
  mutate(link = paste0("https://macaulaylibrary.org/asset/", 
                       cat_num),
         measurement = "", measurement_units = ""
)

# write the final data set to a csv file
# note: this also checks if you already have a file called "my_data_file.csv" to avoid accidentally overwriting data.
if(file.exists("my_output/my_data_file.csv")==T){warning("my_data_file.csv already exists. Rename it and try again.")} else{
write.csv(my_data_final, "my_output/my_data_file.csv",
          row.names = F)
}

```
# Making your measurements
At this point, minimize RStudio, go back to the README file, and complete steps 9-14. In those steps, you'll use spreadsheet software to open the "my_data_file.csv" file you just created. Then, you'll go to each link, view the sonogram, and make the measurements you came up with using a ruler or visual observation. (See README file for more info.) You'll add those measurements to your data file and then come back here to graph the results.

# Graph the results
Use this code to plot your results. It will save a png file to the "my_output" folder. Before you run it, change the 2nd and 3rd lines of the code to reflect your correct units and species. After you run this chunk of code, go ahead and complete the remaining sections of code in this file (described in README steps 15-17).
```{r}
# input the correct measurement units and species
your_units <- "Number of syllables in trill"
your_species <- "Prairie Warbler"

# read in the completed data file
data_complete <- read.csv("my_output/my_complete_data_file.csv")

# plot the data
data_complete %>%
  ggplot(aes(x = Latitude, y = measurement)) +
  geom_point(size = 4) +
  geom_smooth(method = "loess", se = F, color = "firebrick") +
  geom_smooth(method = "lm", se = T, color = "darkgray", lty = 2) +
  theme_bw() + 
  theme(text = element_text(size = 14)) +
  labs(y = your_units, title = your_species)

# save the plot
ggsave("my_output/my_plot.png", height = 4, width = 6, dpi = 300)

```
# Describe your methods
Replace the DESCRIBE YOUR METHODS HERE text with your own. Don't delete any of the quotation marks or other text. Be sure to clearly describe what you measured and how. Write as much as you need to, but only replace the text within the asterisks(*). After you are done writing, run the code to write a text file using the green arrow.
```{r}
cat(file = "my_output/my_methods.txt", 
    "
METHODS

*DESCRIBE YOUR METHODS HERE*


    "
)

```
# Describe your results
Replace the DESCRIBE YOUR RESULTS HERE text with your own. Don't delete any of the quotation marks or other text. Write as much as you feel you need to, but only replace the text within the asterisks(*). After you are done writing, run the code to write a text file using the green arrow.
```{r}
cat(file = "my_output/my_results.txt", 
    "
RESULTS

*DESCRIBE YOUR RESULTS HERE*


    "
)
```
